The Forward March of Categorical Tolerance in the United States\thanks{Reserved for acknowledgments}

### 2025-11-15 manuscript.tex

\newpage
\begin{abstract}
This paper updates the empirical picture of Categorical Tolerance (CT), namely, the pattern of refusing to report dislikes across cultural genres, in the United States during the third decade of the twenty-first century. Analyzing recent survey data from two platforms, I find that CT has continued its march among Americans, reaching approximately 20\% of respondents. The analysis confirms earlier-observed demographic trends, showing that CT is strongly associated with younger cohorts and non-white individuals. However, I also find that individuals reporting the highest educational attainment are now overrepresented among categorical tolerants, suggesting that CT may increasingly function as an elite cultural strategy consistent with contemporary forms of status display, signaling openness and refusal to refuse. Furthermore, the study uses political ideology as a predictor of symbolic exclusion, revealing that while the odds of being a CT itself are not strongly polarized, the inclination toward symbolic exclusion among non-CTs is, with conservatives significantly more likely to express a greater volume of cultural dislikes than liberals. This result shows that political ideology, alongside age and race, is a central determinant of cultural exclusion in the contemporary context. 
\end{abstract}
\newpage
\section*{Introduction}
Ever since Richard Peterson introduced the felicitous term ``patterns of cultural choice'' into sociological discourse more than four decades ago \citep{peterson1983patterns-f85}, sociologists have been busy empirically uncovering and debating the sociological reality, drivers, and correlates of various such patterns. These include, most famously and influentially, the Petersonian cultural \textit{omnivore} \citep{peterson1996changing-967, goldberg2011mapping-77a, warde2008omnivorous-e77} along with their evil twin the culturally exclusive \textit{univore} \citep{bryson1997what-933, peterson1992understanding-ce3, tonkovi2025globallocal-96f}, but also culturally \textit{voracious} consumers \citep{sullivan2007omnivore-018}, \textit{poly-purists} and \textit{poly-mixers} \citep{goldberg2016what-4e3, poecke2018pure-873}, \textit{decisives} and \textit{indecisives} \citep{sonnett2004musical-961}, and even plain old \textit{snobs} engaging in (presumably) Bourdieusian boundary-drawing distinction \citep{atkinson2011context-d44, nault2021social-1bf}, among many others.

In a paper published almost a decade ago, \citet{lizardo2016end-4fb}---hereafter L\&S---introduced another such pattern of cultural choice, which they labeled \textit{categorical tolerance}, to refer to the phenomenon of a growing proportion of the population who refuse to report dislikes of \textit{any} cultural genres---typically musical genres---in standard social science surveys designed to measure patterns of cultural taste. In that paper, L\&S noted that categorical tolerance appeared to be on the rise in the United States, going from about five percent of the population in the canonical 1993 General Social Survey culture module dataset to about sixteen percent in a then-recent survey of a representative sample of Americans collected in 2012, which matched the GSS 1993 battery of items on likes and dislikes of musical tastes across eighteen genres. This resulted in a significant shift in the shape of the univariate count distributions of number of dislikes, which went from a standard Poisson distribution in the GSS 1993 to a count distribution with excess zeroes in 2012 \citep[p.90, Figure 1]{lizardo2016end-4fb}.

In the same paper, L\&S also showed that the rise of categorical tolerance among Americans was a socially patterned phenomenon. Particularly, L\&S showed that refusal to dislike was a growing trend among more recently born cohorts, suggestive of a macro-level pattern of cultural change via cohort replacement of the type that has been more recently shown to govern macro-level cultural change across a wide variety of values, attitudes, beliefs, and practices \citep{vaisey2016cultural-867, kiley2020measuring-123, ochoa2024opinions-856, vaisey2021model-based-deb}. Moreover, when comparing 2012 to 1993 data, L\&S found that categorical tolerants were over-represented among non-white respondents, with a slight under-representation of people with a college degree, indicating that categorical tolerance did not seem to be driven by the usual status and cultural-capital-based dynamics emphasized by Bourdieusian models of social distinction \citep{bourdieu1984distinction-835}, which usually focuses on status display among higher educated elites \citep{jarness2017im-001, khan2011privilege-68a}.

This paper uses recently collected data from a representative sample of the American population to update the empirical picture of categorical tolerance in the United States for the third decade of the twenty-first century. First, I ask whether we can observe a continued rise in the categorical tolerance effect in the more recent data. Secondly, I examine the socio-demographic correlates of categorical tolerance to determine whether this pattern of cultural choice persists in a similar socio-demographic ``niche'' \citep{mcpherson2004blau-9fe} as L\&S observed in 2012, or whether its location in social space has evolved since then. Third, I also examine the socio-demographic correlates of \textit{symbolic exclusion} among people at risk of expressing dislikes \citep{bryson1996anything-311}, to see if we can observe similar patterns to those previously uncovered by L\&S. Finally, the paper adds a missing dimension to previous work on categorical tolerance and symbolic exclusion by investigating whether these cultural taste patterns are connected to political identification as liberal or conservative, given recent work indicating that cultural taste patterns have themselves become embroiled in the ``oil spill'' produced by the continuing political polarization of the American mass public in the last two decades. This phenomenon has accelerated in the Trump era \citep{dellaposta2020pluralistic-ed4, rawlings2023polarization-0af}.

To preview the main results: I find that categorical tolerance has indeed continued its onward march among Americans, with about 20\% of respondents failing to report disliking any of the 12 genres in data collected from two survey samples on two separate platforms. Secondly, the social location of categorical tolerance is fairly similar to that observed by L\&S a decade ago; younger adults are overrepresented among categorical tolerants relative to older adults. Similarly, as L\&S found in 2012, non-white respondents are overrepresented among categorical tolerants in the United States. One distinct difference between the current findings and L\&S's results, however, is a seeming trickle-up of categorical tolerance as an elite taste pattern. Those reporting the highest levels of education are overrepresented in this group, indicating that Bourdieusian status display dynamics may increasingly be at play for this pattern of taste. Finally, while categorical tolerance is not polarized along liberal/conservative lines, \textit{symbolic exclusion} patterns net of categorical tolerance are. After adjusting for other socio-demographic predictors, conservatives are more likely to express a greater number of cultural dislikes than liberals. This indicates that political ideology, along with age---but not education---is now the most reliable predictor of expressing dislikes for cultural genres in the United States today.

@@

The general empirical implication is that cultural omnivores should be more (explicitly) tolerant, liberal, and inclusive, displaying a fundamental openness to consume and engage in a wide variety of cultural practices and experiences. However, the Petersonian thesis does not imply that omnivores indiscriminately accept all cultural forms. Instead, subsequent research goes through some pains to emphasize that cultural omnivores still engage in ``patterned tolerance" and selective exclusion of low-status cultural forms \citep{bryson1996anything-311, lindblom2019anything-530}. This means that while cultural elites embrace a wide variety of cultural genres, they often express explicit dislikes toward a few specific genres, particularly those associated with lower-status groups \citep{bryson1996anything-311, lindblom2022growing-ded}.

Why are dislikes expressed in social surveys on cultural tastes important? As \citet{bourdieu1984distinction-835} himself noted, taste is often asserted negatively through refusal (p. 56). \citet{bryson1996anything-311} linked this dynamic to the idea of ``symbolic exclusion'' or the \textit{indirect} action of drawing a symbolic (social) boundary against particular (usually lower status) groups by distancing oneself from the cultural objects that those groups are perceived to consume \citep{lizardo2016cultural-aaa}. This recognition has led scholars to stress the critical importance of studying ``dislikes" alongside ``likes" to fully comprehend the dynamics of symbolic exclusion and the construction of cultural boundaries \citep{lizardo2015musical-8c6, lindblom2022growing-ded, oncini2023cultural-323}. This general framework has proven helpful for understanding social and cultural dynamics, even beyond the United States. For example, a recent study of musical taste conducted in South Korea found that tolerant omnivores still excluded genres like ``gugak" (Korean traditional music) and ``trot," (one of the oldest forms of Korean popular music) which are preferred by lower-educated and older individuals, suggesting that the more ``cosmopolitan'' omnivores still engage in patterns of symbolic exclusion directed at older persons of lower educational status \citep[321]{kim2020anything-71b}.

@@

Various studies conducted since Bryson's consistently show an empirical link between cultural openness and social tolerance across multiple contexts. For instance, \citet{cho2023are-61e} show that cultural omnivores in South Korea are more accepting of immigrants, demonstrating a greater willingness to include them as friends, neighbors, and even marriage partners. Beyond immigration, cultural omnivorousness is positively correlated with a liberal political outlook and a higher degree of social trust, particularly towards outgroup members. Their diverse cultural engagements are also associated with a cosmopolitan outlook, support for social integration and inclusiveness, and progressive values such as the rejection of traditional gender roles and acceptance of alternative sexual lifestyles \citep{chan2019understanding-1d0}. In line with this work, A study conducted in Finland by \citep{lindblom2022growing-ded} indicates that omnivores hold very liberal views on immigration, law enforcement, and same-sex relationships, while prioritizing environmental protection \citep{lindblom2022growing-ded}.

Conversely, Univorism, a concept first conceptualized by \citet{peterson1992understanding-ce3} and empirically studied by \citet{bryson1997what-933}, describes individuals characterized by narrow, limited cultural tastes, typically confining themselves to a narrower range of popular cultural practices \citep{lindblom2022growing-ded, cho2023are-61e}. This restricted taste profile is often characteristic of older age groups and those with lower educational attainment \citep{sokolova2020does-65c, lindblom2019anything-530}. Recent work shows that univores exhibit less liberal and less tolerant attitudes across various social and political domains. For instance, univores tend to be positioned on the conservative end of the liberal-conservative spectrum, are more skeptical towards fellow citizens, indicating lower social trust,  have lower voter turnout, and espouse more conservative political views that are critical of immigration and generally do not prioritize environmental protection \citep{lindblom2022growing-ded, cho2023are-61e, chan2019understanding-1d0}.

\subsection*{The Rise of Categorical Tolerance}
The concept of categorical tolerance (CT), as introduced by L\&S, identifies a cultural taste orientation characterized by an individual's refusal to express dislikes toward any cultural genre or item. In this way, CT effectively translates into a pattern of exhibiting no explicit symbolic exclusion \citep{bryson1996anything-311}, a taste pattern traditionally linked to individuals of high social status, implying a form of ``openness to cultural diversity" that presumably serves as a new ethos for privileged strata \citep{ollivier2008modes-96f, lindblom2019anything-530}. Conceptually, however, CT can be positioned as a distinct taste orientation alongside cultural omnivorousness, which involves a wide range of ``likes" and very few ``dislikes'', and univorousness, characterized by a narrow range of ``likes" and numerous ``dislikes'' \citep{lindblom2022growing-ded, sonnett2016ambivalence-352}. This orientation suggests a potential shift in how individuals navigate cultural boundaries and express particular forms of taste in the current setting \citep{lizardo2016end-4fb}

In their original research using data from the United States, L\&S found that the phenomenon of categorical tolerance has notably increased over time among non-white and non-elite populations. Recent work by \citet{lindblom2022growing-ded} most closely follows L\&S's original design, namely, comparing two repeated cross-sections of the same population. Using a survey instrument querying respondents on their likes and dislikes for thirteen musical genres, \citet{lindblom2022growing-ded} finds that the proportion of categorical tolerants (persons who refuse to dislike any of the genres) increased in Finland from about fourteen percent of the population in 2007 to twenty percent in 2018. Regarding the social distribution of categorical tolerance, aside from women being over-represented, \citet{lindblom2022growing-ded} found few socio-demographic correlates but observed that CTs displayed greater ambivalence in their taste preferences (e.g., ``mixed feelings'') compared to both omnivores and univores. Their social indistinctiveness also increased over time \citep{lindblom2022growing-ded}. This suggests that for CTs, the absence of dislikes might stem from indecisiveness or a reluctance to take a definitive stance rather than from a genuine embrace of distinct cultural forms \citep{lindblom2022growing-ded}. In the end, \citet[p. 13]{lindblom2022growing-ded} concluded that the taste orientation signaled by categorical tolerance, while on the rise, perhaps signaled a kind of social unstructured ambivalence \citep{sonnett2004musical-961}. However, unlike L\&S's study, \citet{lindblom2022growing-ded} did not use a statistical approach (e.g., zero-inflated count models) to separate respondents at risk of expressing dislikes from those who are true refusers. As such, whether categorical tolerance has a coherent social niche in the contemporary context remains somewhat of an open question.

Subsequent work has extended the CT framework to study other forms of cultural taste beyond music. For instance, in the realm of culinary tastes among young urban Finns, a significant segment of CTs was identified. While these individuals exhibited fewer overall dislikes, higher-status groups, particularly those with higher education, still demonstrated ``patterned intolerance" \citep{bryson1996anything-311}, by selectively disliking certain ``fast and convenient" cuisines, reinforcing symbolic boundaries \citep{lindblom2019anything-530}. Conversely, among less educated groups, categorical tolerance in food preferences was sometimes interpreted as a form of ``cultural goodwill" \citep{bourdieu1984distinction-835}, or an emulation of socially valued tastes on the part of middle-status persons \citep{lindblom2019anything-530}. Additionally, in a study of food and drink avoidances in Italy \citep{oncini2023cultural-323}, a CT cluster was also identified (e.g., expressing little to no culinary dislikes), whose members avoided fewer items than average, resembled cultural omnivores, and were more likely to be younger, male, and possessing higher cultural and economic capital \citep{oncini2023cultural-323}. These findings highlight that the manifestation and social correlates of categorical tolerance can vary, reflecting complex interactions between individual dispositions, socioeconomic factors, and specific cultural contexts \citep{lindblom2019anything-530, oncini2023cultural-323}.

In sum, the recent literature on cultural taste has converged around two main themes. One strand of work, building on the foundational work by \citep{bryson1996anything-311}, has explored the connection, perhaps rising in recent times, between ``cultural tolerance'' (as expressed in the refusal to dislike other cultural forms) and various forms of attitudinal tolerance and ``liberalism'' across various socio-political domains. This work suggests that in the context of increasing political polarization and even rising political ``sectarianism" \citep{finkel2020political-78d}, we should observe a linkage between cultural taste and political ideology \citep{rawlings2023polarization-0af}, with liberals being less likely to engage in symbolic exclusion than conservatives. Secondly, work on ``categorical tolerance'' suggests that this particular pattern of taste, characterized by the universal refusal to engage in \textit{any} form of symbolic exclusion may be on the rise, and could become characteristic of a particular (younger, less white) segment of the population in the United States; alternatively, as it has spread, categorical tolerance may have lost its social moorings, not being strongly correlated with any markers of social position \citep{lindblom2022growing-ded}. By the same token, if categorical tolerance is part of the increasingly observed package of ``conspicuous openness to diversity'' evident in the avowed tastes of contemporary elites \citep{ollivier2008modes-96f}, then it may have even undergone an upgrading in the social space, being more likely to be espoused by individuals with high cultural capital. The empirical analysis that follows addresses both of these themes in the context of the second decade of the twenty-first century.

\section*{Data}
The data for this study come from three rounds of survey data collection on Americans' cultural tastes. The two waves of data collection yielded two convenience samples: one collected in 2021 and the other in 2023, with an oversample of people with advanced educational degrees. Both of these samples come from the Prolific Academic (ProA) platform. Prolific has emerged as a preferred alternative to Amazon’s Mechanical Turk for researchers seeking high-quality, suitably motivated participant pools. Unlike MTurk, ProA was explicitly designed for social-scientific researchers and offers a high-quality, better-compensated participant pool. Also, unlike MTurk, ProA mandates a minimum payment based on the expected time required to complete the task. Participants in our study were compensated at a rate of \$9.50 per hour, which ProA indicates as a “good” compensation rate. The main participant pool for this study on the Prolific platform was limited to participants between the ages of eighteen and eighty-five who reside in the United States and whose first language is English (or are fluent in English). The study was also restricted to participants who scored at or above 95\% on ProA’s ``approval rate'' (the proportion of times a participant’s study response has been accepted in other studies as adequate). Survey completion was verified by embedding a completion code as the last question, which participants had to input on the Prolific website to confirm they had read and completed each survey question.

In the summer of 2025, I collected a third sample to replicate the results of the first two waves using a more representative sample of American respondents. For this purpose, I used the recently launched Cint (previously Lucid) Theorem platform, which provides access to a quota-sampled set of respondents matched to the U.S. census by age, gender, ethnicity, and region. Like Prolific, Cint Theorem is designed primarily for social science research. Cint employs a "marketplace" approach to recruiting study participants by connecting researchers with various suppliers and offering a variety of incentives to respondents. In this way, Cint provides relatively affordable access to participants for researchers (about \$1.50 per completed survey) and high-quality samples \citep{crabtree2025yes-e5f}. Cint also uses various automated screens (e.g., invalid traffic, duplicate IPs) and response quality indicators (e.g., CAPTCHA scores) to weed out bots and prevent less attentive participants from accessing the survey. Representative samples of the U.S. population from Cint Theorem have been used effectively in recent social scientific work on Americans' political and social attitudes (e.g., Keskintürk et al. 2025).

@@

\subsection*{Univariate Distribution of Number of Dislikes}
Table~\ref{tab:dis-dist} shows the univariate distribution of the number of dislikes for both the Prolific and Lucid samples. As we can see, both datasets closely agree on the proportion of categorical tolerants, with the Prolific data yielding about 19\% and the Lucid dataset about 20\%. When we combine both, we get an estimate of 19.6\%, suggesting that about a fifth of the American population today can be classified as categorical tolerants, a presumed increase from L\&S's estimate of about 15\% using data from 2012.\footnote{We should be cautious when making direct comparisons, given that the number of genres included is smaller (twelve versus twenty in L\&S's 2012 survey) in both the Prolific and Lucid samples. Nevertheless, we can conclude that categorical tolerance remains a pervasive phenomenon among a substantial number of Americans.}

@@

As shown in Figure~\ref{fig:main}, the distribution of the number of dislikes we get from the joint samples (like the separate sample distributions), exhibits excess zeroes, similar to the one obtained by L\&S using 2012 data. Even considering issues of sample comparability, we can conclude that the taste regime governing patterns of cultural choice among Americans in the second decade of the twenty-first century certainly no longer resembles that captured by the GSS in the early 1990s, with a larger-than-expected---if the number of dislikes followed a simple Poisson process---number of people refusing to express dislikes for any genre.

>>>>

\newpage
\begin{abstract}
This paper updates the empirical picture of Categorical Tolerance (CT), namely, the pattern of refusing to report dislikes across cultural genres, for the third decade of the twenty-first century in the United States. Analyzing recent survey data from two platforms, I find that CT has continued its march among Americans, reaching approximately 20\% of respondents. The analysis confirms earlier-observed demographic trends, showing that CT is strongly associated with younger cohorts and non-white individuals. However, I also find that individuals reporting the highest educational attainment are now overrepresented among categorical tolerants, suggesting that CT may increasingly function as an elite cultural strategy consistent with contemporary forms of status display, signaling openness and refusal to refuse. Furthermore, the study uses political ideology as a predictor of symbolic exclusion, revealing that while the odds of being a CT itself are not strongly polarized, the inclination toward symbolic exclusion among non-CTs is, with conservatives significantly more likely to express a greater volume of cultural dislikes than liberals, showing that political ideology, alongside age and ethnoracial identity, is a central determinant of cultural exclusion in the contemporary context. 
\end{abstract}
\newpage
\section*{Introduction}
Ever since Richard Peterson introduced the felicitous term ``patterns of cultural choice'' into sociological discourse more than four decades ago \citep{peterson1983patterns-f85}, sociologists have been busy empirically uncovering and debating the sociological reality, drivers, meaning, and correlates of various such patterns. These include, most famously and influentially, the Petersonian cultural \textit{omnivore} \citep{peterson1996changing-967, goldberg2011mapping-77a, warde2008omnivorous-e77} along with their evil twin the culturally exclusive \textit{univore} \citep{bryson1997what-933, peterson1992understanding-ce3, tonkovi2025globallocal-96f}, but also culturally \textit{voracious} consumers \citep{sullivan2007omnivore-018}, \textit{poly-purists} and \textit{poly-mixers} \citep{goldberg2016what-4e3, poecke2018pure-873}, \textit{decisives} and \textit{indecisives} \citep{sonnett2004musical-961}, and even plain old \textit{snobs} engaging in (presumably) Bourdieusian boundary-drawing distinction \citep{atkinson2011context-d44, nault2021social-1bf}, among many others.

In a paper published almost a decade ago, \citet{lizardo2016end-4fb}---hereafter L\&S---introduced another such pattern into sociological discourse, which they labeled \textit{categorical tolerance}, to refer to the phenomenon of a growing proportion of the population who refuse to report dislikes of \textit{any} cultural genres---typically musical genres---in standard social science surveys designed to measure taste and consumption practices. In that paper, L\&S noted that categorical tolerance appeared to be on the rise in the United States, going from about five percent of the population in the canonical 1993 General Social Survey culture module dataset to about sixteen percent in a then-recent survey of a representative sample of Americans collected in 2012, which matched the GSS 1993 battery of items on likes and dislikes of musical tastes across eighteen genres. This resulted in a significant shift in the shape of the univariate count distributions of number of dislikes, which went from a standard Poisson distribution in the GSS 1993 to a count distribution with excess zeroes in 2012 \citep[p.90, Figure 1]{lizardo2016end-4fb}.

In the same paper, L\&S also showed that the rise of categorical tolerance among Americans was a socially patterned phenomenon. Particularly, L\&S showed that refusal to dislike was a growing trend among more recently born cohorts, suggestive of a macro-level process of cultural change via cohort replacement of the type that has been more recently shown to govern macro-level cultural change across a wide variety of values, attitudes, beliefs, and practices \citep{vaisey2016cultural-867, kiley2020measuring-123, ochoa2024opinions-856, vaisey2021model-based-deb}. Moreover, when comparing 2012 to 1993 data, L\&S found that categorical tolerants were over-represented among non-white respondents, with a slight under-representation of people with a college degree, indicating that categorical tolerance did not seem to be driven by the usual status and cultural-capital-based dynamics emphasized by Bourdieusian models of social distinction \citep{bourdieu1984distinction-835}, which usually focuses on status display among higher educated elites \citep{jarness2017im-001, khan2011privilege-68a}.

This paper uses recently collected data from a representative sample of the American population to update the empirical picture of categorical tolerance in the United States for the third decade of the twenty-first century. First, I ask whether we can observe a continued rise in the categorical tolerance effect in the more recent data. Secondly, I examine the socio-demographic correlates of categorical tolerance to determine whether this pattern of cultural choice persists in a similar socio-demographic ``niche'' \citep{mcpherson2004blau-9fe} as L\&S observed in 2012, or whether its location in social space has evolved since then. Third, I also examine the socio-demographic correlates of \textit{symbolic exclusion} among people at risk of expressing dislikes \citep{bryson1996anything-311}, to see if we can observe similar patterns to those previously uncovered by L\&S. Finally, the paper adds a missing dimension to previous work on categorical tolerance and symbolic exclusion by examining whether these cultural taste patterns are connected to political identification as liberal or conservative, given recent work indicating that cultural taste patterns have themselves become embroiled in the ``oil spill'' \citep{dellaposta2020pluralistic-ed4} produced by the continuing political polarization of the American mass public in the last two decades, a phenomenon that has accelerated in the Trump era \citep{rawlings2023polarization-0af}.

To preview the main results: I find that categorical tolerance has indeed continued its onward march among Americans, with about 20\% of respondents failing to report disliking any of the 12 genres in data collected from three survey samples on two separate platforms. Secondly, the social location of categorical tolerance is fairly similar to that observed by L\&S a decade ago; younger adults are overrepresented among categorical tolerants relative to older adults. Similarly, as L\&S found in 2012, non-white respondents are overrepresented among categorical tolerants in the United States. One distinct difference between the current findings and L\&S's results, however, is a seeming trickle-up of categorical tolerance as an elite taste pattern. Those reporting the highest levels of education are overrepresented in this group, indicating that Bourdieusian status display dynamics may increasingly be at play. Finally, while categorical tolerance is not polarized along liberal/conservative lines, \textit{symbolic exclusion} patterns net of categorical tolerance are. After adjusting for other socio-demographic predictors, conservatives are more likely to express a greater number of cultural dislikes than liberals. This indicates that political ideology, along with age---but not education---is now the most reliable predictor of expressing dislikes for cultural genres in the United States today.

@@

The general empirical implication is that cultural omnivores should be more (explicitly) tolerant, liberal, and inclusive, displaying a fundamental openness to consume and engage in a wide variety of cultural practices and experiences. However, the Petersonian thesis does not imply that omnivores indiscriminately accept all cultural forms. Instead, subsequent research goes to some lengths to emphasize that cultural omnivores still engage in ``patterned tolerance" and the selective exclusion of low-status cultural forms \citep{bryson1996anything-311, lindblom2019anything-530, warde2008omnivorous-e77, prieur2008cultural-0db}. This means that while cultural elites embrace a wide variety of cultural genres, they often express explicit dislikes toward and symbolic distancing from a few specific genres, particularly those associated with lower-status groups \citep{bryson1996anything-311, lindblom2022growing-ded}.\footnote{This last is what \citet{vries2022what-1a2} call the ``weak'' omnivore thesis, but which like many things weak in sociology, it is also the only empirically plausible version of the thesis.}

Why are dislikes expressed in social surveys on cultural tastes important? As \citet[56]{bourdieu1984distinction-835} himself noted, taste is often asserted negatively via refusal. \citet{bryson1996anything-311} linked this dynamic to the idea of ``symbolic exclusion'' or the \textit{indirect} action of drawing a symbolic (social) boundary against particular (usually lower status) groups by distancing oneself from the cultural objects that those groups are perceived to consume \citep{lizardo2016cultural-aaa}. This recognition has led scholars to stress the critical importance of studying ``dislikes" alongside ``likes" to fully understand the dynamics of symbolic exclusion and the construction of cultural boundaries \citep{lizardo2015musical-8c6, lindblom2022growing-ded, oncini2023cultural-323}. Importantly, this general framework has proven helpful for understanding social and cultural dynamics, beyond the case of the United States. For example, a recent study of musical taste conducted in South Korea found that tolerant omnivores still excluded genres like ``gugak" (Korean traditional music) and ``trot," (one of the oldest forms of Korean popular music) which are preferred by lower-educated and older individuals, suggesting that the more ``cosmopolitan'' omnivores still engage in patterns of symbolic exclusion directed at older persons of lower educational status \citep[321]{kim2020anything-71b}.

@@

Various studies conducted since Bryson's consistently show an empirical link between cultural openness and social tolerance across multiple contexts. For instance, \citet{cho2023are-61e} show that cultural omnivores in South Korea are more accepting of immigrants, demonstrating a greater willingness to include them as friends, neighbors, and even marriage partners. Beyond immigration, cultural omnivorousness is positively correlated with a liberal political outlook and a higher degree of social trust, particularly towards outgroup members. Cultural omnivorousness is also associated with cosmopolitan values, support for social integration and inclusiveness, and progressive stances such as the rejection of traditional gender roles and acceptance of alternative sexual lifestyles \citep{chan2019understanding-1d0}. In line with this work, A study conducted in Finland by \citet{lindblom2022growing-ded} indicates that omnivores are more likely to hold liberal views on immigration, law enforcement, and same-sex relationships, while also prioritizing environmental protection \citep{lindblom2022growing-ded}.

Conversely, Univorism, a concept first conceptualized by \citet{peterson1992understanding-ce3} and empirically studied by \citet{bryson1997what-933}, describes individuals characterized by narrow, limited cultural tastes, typically confining themselves to a more delimited range of popular cultural practices \citep{lindblom2022growing-ded, cho2023are-61e}. This restricted taste profile is often characteristic of older age groups and those with lower educational attainment \citep{sokolova2020does-65c, lindblom2019anything-530}. Recent work shows that univores exhibit less liberal and less tolerant attitudes across various social and political domains. For instance, univores tend to be positioned on the conservative end of the liberal-conservative spectrum, are more skeptical towards fellow citizens, indicating lower social trust,  have lower voter turnout, are more critical of immigration, and generally do not prioritize environmental protection \citep{lindblom2022growing-ded, cho2023are-61e, chan2019understanding-1d0}.

\subsection*{The Rise of Categorical Tolerance}
The concept of categorical tolerance (CT), as introduced by L\&S, identifies a cultural taste orientation characterized by an individual's refusal to express dislikes toward any cultural genre or item. In this way, CT effectively translates into the refusal to show explicit forms of symbolic exclusion \citep{bryson1996anything-311}. As noted, a lower likelihood to engage in symbolic exclusion via cultural dislikes is a taste pattern traditionally linked to individuals of high social status, implying a form of ``openness to cultural diversity" that presumably serves as a new ethos for privileged strata \citep{ollivier2008modes-96f, lindblom2019anything-530}. Conceptually, however, CT can be positioned as a distinct taste orientation relative to cultural omnivorousness, which involves a wide range of ``likes" and very few (but not the absence of) ``dislikes'', and univorousness, characterized by a narrow range of ``likes" and numerous ``dislikes'' \citep{sonnett2004musical-961}. This orientation suggests a potential shift in how individuals navigate cultural boundaries and express particular forms of taste in the current setting \citep{lizardo2016end-4fb}

In their original research using U.S. data, L\&S found that the phenomenon of categorical tolerance has notably increased over time among non-white and non-elite populations. Recent work by \citet{lindblom2022growing-ded} most closely follows L\&S's original design, namely, comparing two repeated cross-sections of the same population. Using a survey instrument querying respondents on their likes and dislikes for thirteen musical genres, \citet{lindblom2022growing-ded} finds that the proportion of categorical tolerants (persons who refuse to dislike any of the genres) increased in Finland from about fourteen percent of the population in 2007 to twenty percent in 2018. Regarding the social distribution of categorical tolerance, aside from women being over-represented, \citet{lindblom2022growing-ded} found few socio-demographic correlates but observed that CTs displayed greater ambivalence in their taste preferences (e.g., ``mixed feelings'') compared to both omnivores and univores, with social non-distinctiveness also increasing over time \citep{lindblom2022growing-ded}.

This suggests that for CTs, the absence of dislikes might stem from indecisiveness or a reluctance to take a definitive stance rather than from a genuine embrace of distinct cultural forms \citep{lindblom2022growing-ded}. In the end, \citet[p. 13]{lindblom2022growing-ded} concluded that the taste orientation signaled by categorical tolerance, while on the rise, perhaps signaled a kind of social unstructured ``ambivalence'' \citep{sonnett2004musical-961}. However, unlike L\&S's study, \citet{lindblom2022growing-ded} did not use a statistical approach (e.g., zero-inflated count models) to distinguish respondents at risk of expressing dislikes from true refusers. As such, whether categorical tolerance has a coherent social niche in the contemporary context remains somewhat of an open question.

Subsequent work has extended the CT framework to study other forms of cultural taste beyond music. For instance, recent work identifies a segment of young urban Finns who exhibit categorical tolerance in culinary tastes. While these individuals exhibited fewer overall dislikes, individuals with higher educational credentials still demonstrated ``patterned tolerance" \citep{bryson1996anything-311}, by selectively disliking certain ``fast and convenient" cuisines, reinforcing class-based symbolic boundaries \citep{lindblom2019anything-530}. Conversely, among less educated groups, categorical tolerance in food preferences was sometimes interpreted as a form of ``cultural goodwill" \citep{bourdieu1984distinction-835}, or an emulation of socially valued tastes on the part of middle-status persons \citep{lindblom2019anything-530}. Additionally, in a study of food and drink avoidances in Italy \citep{oncini2023cultural-323}, a CT cluster was also identified (e.g., expressing little to no culinary dislikes), whose members avoided fewer items than average, resembled cultural omnivores, and were more likely to be younger, male, and possessing higher cultural and economic capital. Overall, these findings highlight that the manifestation and social correlates of categorical tolerance can vary, reflecting complex interactions between individual dispositions, socioeconomic factors, and specific cultural contexts \citep{lindblom2019anything-530, oncini2023cultural-323}.

To summarize, the recent literature on cultural taste has converged around two main themes. One strand of work, building on the foundational work of \citep{bryson1996anything-311}, has explored the connection, perhaps rising in recent times, between ``cultural tolerance'' (as expressed in the refusal to dislike other cultural forms) and various forms of attitudinal tolerance and ``liberalism'' across various socio-political, practical, and attitudinal domains. This work suggests that in the context of increasing political polarization and even rising political ``sectarianism" \citep{finkel2020political-78d}, we should observe a linkage between cultural taste and political ideology \citep{rawlings2023polarization-0af}, with liberals being less likely to engage in symbolic exclusion than conservatives.

Secondly, work on ``categorical tolerance'' suggests that this particular pattern of taste, characterized by the universal refusal to engage in \textit{any} form of symbolic exclusion may be on the rise, and could become characteristic of a particular (younger, less white) segment of the population in the United States; alternatively, as it has spread, categorical tolerance may have lost its social moorings, not being strongly correlated with any markers of social position \citep{lindblom2022growing-ded}. By the same token, if categorical tolerance is part of the increasingly observed package of ``conspicuous openness to diversity'' evident in the avowed tastes of contemporary elites \citep{ollivier2008modes-96f}, then it may have even undergone an upgrading in the social space, being more likely to be espoused by individuals with high cultural capital. The empirical analysis that follows addresses both themes in the context of the third decade of the twenty-first century in the U.S.

\section*{Data}
The data for this study come from three rounds of survey data collection on Americans' cultural tastes. The two waves of data collection yielded two convenience samples: one collected in 2021 and the other in 2023, the latter with an oversample of people with advanced educational degrees. Both of these samples come from the Prolific Academic (ProA) platform. Prolific has emerged as a preferred alternative to Amazon’s Mechanical Turk for researchers seeking high-quality, suitably motivated participant pools. Unlike MTurk, ProA was explicitly designed for social-scientific researchers and offers a high-quality, better-compensated participant pool. Also, unlike MTurk, ProA mandates a minimum payment based on the expected time required to complete the task. Participants in our study were compensated at a rate of \$9.50 per hour, which ProA indicates as a “good” compensation rate. The main participant pool for this study on the Prolific platform was limited to participants between the ages of eighteen and eighty-five who reside in the United States and whose first language is English (or are fluent in English). The study was also restricted to participants who scored at or above 95\% on ProA’s ``approval rate'' (the proportion of times a participant’s study response has been accepted in other studies as adequate). Survey completion was verified by embedding a completion code as the last question, which participants had to input on the Prolific website to confirm they had read and completed each survey question.

In the summer of 2025, a third sample was collected to replicate the results of the first two waves using a more representative sample of American respondents, using the relatively recently launched Cint (previously Lucid) Theorem platform, which provides access to a quota-sampled set of respondents matched to the U.S. census by age, gender, ethnicity, and region. Like Prolific, Cint Theorem is designed primarily for social science research. Cint employs a "marketplace" approach to recruiting study participants by connecting researchers with various suppliers and offering a variety of incentives to respondents. In this way, Cint provides relatively affordable access to participants for researchers (about \$1.50 per completed survey) and high-quality samples \citep{crabtree2025yes-e5f}. Cint also uses various automated screens (e.g., invalid traffic, duplicate IPs) and response quality indicators (e.g., CAPTCHA scores) to weed out bots and prevent less attentive participants from accessing the survey. Representative samples of the U.S. population from Cint Theorem have been used effectively in recent social scientific work on Americans' political and social attitudes (e.g., Keskintürk et al. 2025).

@@

\subsection*{Univariate Distribution of Number of Dislikes}
Table~\ref{tab:dis-dist} shows the univariate distribution of the number of dislikes for the Prolific and Lucid samples. As we can see, both datasets closely agree on the proportion of (potential) categorical tolerants, with the Prolific data yielding about 19\% and the Lucid dataset about 20\%. When we combine both, we get an estimate of 19.6\%, suggesting that about a fifth of the American population today may be classified as categorical tolerants, a presumed increase from L\&S's estimate of about 15\% using 2012 data.

@@

As shown in Figure~\ref{fig:main}, the distribution of the number of dislikes we get from the pooled samples (like the separate sample distributions), exhibits excess zeroes, similar to the one obtained by L\&S using 2012 data. Even considering issues of sample comparability relative to L\&S's original study, we can conclude that the taste regime governing patterns of cultural choice among Americans in the second decade of the twenty-first century certainly no longer resembles that captured by the GSS in the early 1990s, with a larger-than-expected---if the number of dislikes followed a simple Poisson process---number of people refusing to express dislikes for any genre.
